Peer-to-Peer Permissionless Consensus via Reputation

### 2023-06-08 freechains/freechains.tex
...

Content publishing in Internet forums and social media is
increasingly more centralized in a few companies (e.g. Facebook and
Twitter)~\cite{internet.fixing}.
%,p2p.osn,p2p.dosn}.
%These companies benefit from closed protocols and network effects to keep
%users locked in their platforms.
On the one hand, these companies offer free storage, friendly user interfaces,
and robust access.
On the other hand, they concentrate too much power over its users, holding and
monetizing over their private data and behavior.
%
Peer-to-peer alternatives~\cite{p2p.survey} eliminate intermediaries, but
strive to achieve consistency when dealing with malicious users.
%given the decentralization of authority and infrastructure.
% and push to end users the responsibility to manage data and connectivity.

In an ideal Internet forum, all messages or posts
(i)   reach all users,
(ii)  are delivered in a consistent order, and
(iii) are respectful and on topic.
In centralized systems, items (i) and (ii) are trivially achieved assuming
availability and delivery order, while item (iii) assumes that users trust the
service to moderate content.
In decentralized systems, however, none of these demands are easily
accomplished.
A common approach in gossiping protocols is to proactively disseminate posts
among peers until they reach all users~\cite{p2p.survey,p2p.byz}.
However, this approach does not guarantee consensus since posts can be received
in conflicting orders~\cite{p2p.intention}.
%,p2p.dvcs}.
%As an example, antagonistic messages such as \emph{"X is final"} vs
%\emph{"Y is final"} might be sent concurrently, preventing the network to
%determine as a group its intention as \emph{X} or \emph{Y}.
%
Consensus is key to distinguish malicious behavior \emph{at the protocol
level}, and therefore, satisfy item (iii).
Without consensus, each client must rely on local anti-abuse policies, such as
SPAM filters and blacklists, but which only apply a posteriori, when the
abusive messages have already been flooded in the network.
%
%fundamental in collaborative applications, such as open discussion
%forums, chat rooms, and trading services (e.g., delivery \& auction).
%Otherwise, they become impractical due to excess and abuse (e.g., SPAM and
%illicit content).
%At the very core, consensus is

@@

Bitcoin~\cite{p2p.bitcoin} is the first permissionless protocol to resist
Sybil attacks~\cite{p2p.sybil}.
Its key insight is to rely on a scarce resource---the \emph{proof-of-work}---to
establish consensus.
%The protocol maintains a single dispendious timeline consisting of linked
%blocks with transactions, which represent the consensus.
%Alternative timelines need to spend more resources to substitute the consensus.
The protocol is Sybil resistant because it is expensive to write to its
unique timeline (either via proof-of-work or transaction fees).
%
However, Bitcoin and cryptocurrencies in general are not suitable for content
sharing because
    (i)   they enforce a unique timeline to preserve value and immunity to
          attacks;
    (ii)  they lean towards concentration of power due to scaling effects;
    (iii) they impose an external economic cost to use the protocol; and
    (iv)  they rely exclusively on objective rules to reach consensus.
%
These issues threaten our original decentralization goals, and more
importantly, they renounce subjectivity, which is inherent to content
moderation.
%
For instance, a unique timeline implies that all Internet content should be
subject to the same consensus rules, while objective rules alone cannot
distinguish abuse amid content.
%
Another limitation of cryptocurrencies is that it is not possible to revoke
content in the middle of a blockchain, which is inadmissible considering
illicit content (e.g., hate speech).
For instance, Bitcoin transactions permanently store arbitrary human-authored
content, including harmful texts and images~\cite{btc.content}.
%~\cite{btc.data}

In this work, we adapt the idea of scarce resources to reach consensus and
eradicate Sybils, but in the context of social content sharing.
%
Our first contribution is to recognize the actual published contents as the
scarce resources themselves, since they require human work.
%
Work is manifested as new posts, which if approved by others, reward authors
with reputation tokens, which are further used to evaluate other posts with
likes and dislikes.
With such \emph{proof-of-authoring} mechanism, token generation is expensive,
while verification is cheap and made by multiple users.
%
Due to decentralization and network partitions, posts in a timeline form a
causal graph with only partial order, which we promote to a total order based
on the reputation of authors.
The consensus order is fundamental to detect conflicting operations, such as
likes with insufficient reputation (akin to Bitcoin's double spending).
%
Our second contribution is to allow users to create diversified forums of
interest (instead of a singleton blockchain), each counting as an independent
timeline with its own subjective consensus etiquette.
%
Our third contribution is to support content removal without compromising the
integrity of the decentralized blockchain.
Users have the power to revoke posts with dislikes, and peers are forced to
remove payloads, only forwarding associated blockchain metadata.
%
We integrated the proposed consensus algorithm into \FC~\cite{fcs.sbseg20}, a
practical peer-to-peer (P2P) content dissemination protocol that provides
strong eventual consistency~\cite{p2p.sec}.
%p2p.crdts,
To show the practicability of the consensus mechanism, we simulated months of
activity of a chat channel and years of a newsgroup forum, both extracted from
publicly available Internet archives.

@@

@@

We propose a reputation system in which users can spend tokens named \reps to
post and rate content in the forums:
a \code{post} initially penalizes authors until it consolidates and counts
positively;
a \code{like} is a positive feedback that helps subscribers to distinguish
content amid excess;
a \code{dislike} is a negative feedback that revokes content when crossing a
threshold.
Table~\ref{fig.general} summarizes the general reputation rules and their
goals.
To prevent Sybils, users with no \reps cannot operate under these rules,
requiring a welcoming like from any other user already in the system.
The fact that likes are zero-sum operations, which only transfer reputation,
eliminates the incentives from malicious users to invite Sybils into the
system.
The only way to generate new \reps is to post content that other users
tolerate, which demands non-trivial work resistant to automation.%
\footnote{
    Note that to some extent AI content may be indistinguishable from human
    content, but its frequency and relevance in forums is still subject to
    human evaluation in our proposal.
}

Bitcoin employs proof-of-work to mitigate Sybil attacks.
However, CPU or other extrinsic resources are not evenly distributed among
humans, specially in communications using battery-powered devices.
%
Instead, considering the context of public forums, we propose to take advantage
of the human authoring ability as an intrinsic resource.
Creating new content is hard and takes time, but is comparatively easy to
verify and rate.
Therefore, in order to impose scarcity, we determine that only content
authoring generates \reps, while likes and dislikes just transfer \reps between
users.
%but limited by periods of time to prevent excess (e.g., once in a day per user).
%Additionally,

@@

@@

Figure~\ref{fig.reps}.A illustrates the evolution of a forum DAG in the
presence of a network partition:
A common prefix has signed posts from users $a$, $b$, and $c$, when they were
still all connected.
%
We assume that within the prefix, users $a$ and $b$ have contributed with
better content and therefore have more reputation combined than $c$ has alone
(i.e., $8+5>3$).
We will use this fact when ordering branches in the consensus algorithm.
%
Then, user $c$ disconnects from $a$ and $b$, and evolves with new users $x$ and
$y$ with posts in \code{branch-1}.
In the meantime, $a$ and $b$ participate with posts in \code{branch-2}.
After some time, the partitions reconnect and new posts from $x$, $b$ and $c$
merge the branches with a common suffix.

Note that even non-malicious network partitions may experience conflicts when
merging branches.
For instance, dislike operations in one branch to common past posts could
possibly block posts from users in the other branch, depending on how they are
finally ordered.
Therefore, we need to order branches with a global consensus, such that we can
validate operations consistently across the network.
%
Note also that, to reach consensus, we should preferably rely on the maximum
information that all branches agree on, which is exactly the posts in the
common prefix.

Therefore, based on the reputation of users in the common prefix,
Figure~\ref{fig.reps}.B indicates the consensus order between posts in the forum
DAG:
    the common prefix comes first,
    then all posts in \code{branch-2},
    then all posts in \code{branch-1},
    then the common suffix.
%
The \code{branch-2} with users $a$ and $b$ takes priority over \code{branch-1}
with user $c$ because, before the forking point, $a$ and $b$ have more
reputation than $c$, $x$, and $y$ combined.

@@

There are some other relevant considerations about forks and merges:
%
Peers that first receive branches with less reputation need to reorder all
posts after the forking point.
This might involve removing content in the end-user software.
This behavior is similar to Bitcoin's blockchain reorganization, when peers
detect new longest chains. % and disconsiders old blocks.
%
Likewise, peers that first saw branches with more reputation just need to
append the other branch, with no reordering at all.
%and do not need to recompute anything.
This behavior is expected to happen in the majority of the connected network.
%
Note that unlike Bitcoin, forks are not only permitted but encouraged due to
the local-first software principle~\cite{p2p.local}, in which networked
applications can work locally while offline.
However, the longer a peer remains disconnected, the more conflicting
operations it may see, and the higher are the chances of reordering when
rejoining.

In summary, it is important to remark that
    (i) branches always introduce conflicts, which
    (ii) require consensus to order operations, which
    (iii) is based on a common prefix reputation,
    (iv) leading to a global and deterministic result.
%
Note that the consensus order exists only to account reputation and verify
operations, and is only a view of the primary DAG structure.

@@

As a counterpoint, maybe users $a$ and $b$ have abandoned the forum for months,
and thus \code{branch-1} is actually legitimate.
In this case, users $a$ and $b$ might be the ones trying to take over the
forum and would succeed, since they control the majority of \reps in the common
prefix.
Yet another possibility is that both branches are legitimate but became
disconnected for a long period of time.
In any case, it is unacceptable for correct users that a very old remote branch
affects a long active forum.

@@

In summary, the rules to merge a branch from a remote machine \code{j} into a
branch from a local machine \code{i} are as follows:
\begin{itemize}
    \item \code{i} is first if it crosses the activity threshold of
          \emph{7 days} or \emph{100 posts}, regardless of \code{j}.
    \item \code{i} or \code{j} is first, whichever has more reputation in the
          common prefix, as discussed in Section~\ref{sec.design.basic}.
    \item otherwise, branches are ordered by an arbitrary criteria, such as
          lexicographical order of the post hashes immediately after the
          common prefix.
\end{itemize}

@@

Because the protocol relies on a tamper-proof DAGs, and because each peer
self verifies the posts as they synchronize, the protocol becomes tolerant to
Byzantine faults~\cite{lamport.byz}.
%
Note that no information coming from any peer is trusted a priori.
%
For instance, a malicious peer that tries to synchronize branches with erratic
data can be detected on the first inconsistent post.
This allows the correct peer to close the connection immediately and blacklist
the malicious peer from future connections.

Therefore, although Byzantine peers may still perform some sort of denial of
service attacks in the network, they cannot modify the DAG structure, create
arbitrary content (without performing work), or impede that two correct nodes
synchronize directly.

@@

\FC~\cite{fcs.sbseg20} is an unstructured P2P topic-based publish-subscribe
protocol, in which each \emph{chain} is a replicated Merkle-DAG representing
the causal relationships between the messages.
%
The protocol operation is typical of publish-subscribe systems: an author
publishes a post to a chain, and subscribed users eventually receive the
message.
% by synchronizing DAGs.

@@

We integrated the proposed reputation and consensus mechanism with \FC to
support content moderation and mitigate abuse.
Table~\ref{fig.rules} details the concrete reputation rules we conceived for
\FC, which is compatible with the general rules of Table~\ref{fig.general}, and
are discussed as follows through an example.
Authors have to sign posts in order to be accounted by the reputation system
and operate in the chains.
We start by creating an identity whose public key is assigned as the pioneer in
a chain named \emph{forum}:

>>>>

Content publishing in Internet forums and social media is
increasingly more centralized in a few companies (e.g. Facebook and
Twitter)~\cite{internet.fixing}.
%,p2p.osn,p2p.dosn}.
%These companies benefit from closed protocols and network effects to keep
%users locked in their platforms.
On the one hand, these companies offer free storage, friendly user interfaces,
and robust access.
On the other hand, they concentrate excessive power, controlling and monetizing
over private user data.
%
Peer-to-peer alternatives~\cite{p2p.survey} eliminate intermediaries, but
strive to achieve consistency when dealing with malicious users.
%given the decentralization of authority and infrastructure.
% and push to end users the responsibility to manage data and connectivity.

In an ideal Internet forum, all messages or posts
(i)   reach all users,
(ii)  are delivered in consistent order, and
(iii) are respectful and on topic.
In centralized systems, items (i) and (ii) are trivially achieved assuming
availability and delivery order, while item (iii) assumes that users trust the
service to moderate content.
In decentralized systems, however, none of these demands are easily
accomplished.
A common approach in gossiping protocols is to proactively disseminate posts
among peers until they reach all users~\cite{p2p.survey,p2p.byz}.
However, this approach does not guarantee consensus since posts can be received
in conflicting orders~\cite{p2p.intention}.
%,p2p.dvcs}.
%As an example, antagonistic messages such as \emph{"X is final"} vs
%\emph{"Y is final"} might be sent concurrently, preventing the network to
%determine as a group its intention as \emph{X} or \emph{Y}.
%
Consensus is key to distinguish malicious behavior \emph{at the protocol
level}, and therefore, satisfy item (iii).
Without consensus, each client must rely on local anti-abuse policies, such as
SPAM filters and blacklists, but which only apply a posteriori, when the
abusive messages have already been flooded in the network.
%
%fundamental in collaborative applications, such as open discussion
%forums, chat rooms, and trading services (e.g., delivery \& auction).
%Otherwise, they become impractical due to excess and abuse (e.g., SPAM and
%illicit content).
%At the very core, consensus is

@@

Bitcoin~\cite{p2p.bitcoin} is the first permissionless protocol to resist
Sybil attacks~\cite{p2p.sybil}.
Its key insight is to rely on a scarce resource---the \emph{proof-of-work}---to
establish consensus.
%The protocol maintains a single dispendious timeline consisting of linked
%blocks with transactions, which represent the consensus.
%Alternative timelines need to spend more resources to substitute the consensus.
The protocol is Sybil resistant because it is expensive to write to its
unique timeline (either via proof-of-work or transaction fees).
%
However, Bitcoin and cryptocurrencies in general are not suitable for content
sharing because
    (i)   they enforce a unique timeline to preserve value and immunity to
          attacks;
    (ii)  they lean towards concentration of power due to scaling effects;
    (iii) they impose an external economic cost to use the protocol; and
    (iv)  they rely exclusively on objective rules to reach consensus.
%
These issues threaten our original decentralization goals, and more
importantly, they renounce subjectivity, which is inherent to content
moderation.
%
For instance, a unique timeline implies that all Internet content should be
subject to the same consensus rules, while objective rules alone cannot
distinguish abuse amid content.
%
Another limitation of cryptocurrencies is that it is not possible to revoke
content in the middle of a blockchain, which is inadmissible considering
illicit content (e.g., hate speech)~\cite{btc.content}.
%For instance, Bitcoin transactions permanently store arbitrary human-authored
%content, including harmful texts and images.
%~\cite{btc.data}

In this work, we adapt the idea of scarce resources to reach consensus and
eradicate Sybils, but in the context of social content sharing.
%
Our first contribution is to recognize the actual published contents as the
scarce resources themselves, since they require human work.
%
Work is manifested as new posts, which if approved by others, reward authors
with reputation tokens, which are further used to evaluate other posts with
likes and dislikes.
With such \emph{proof-of-authoring} mechanism, token generation is expensive,
while verification is cheap and shared by multiple users.
%
Due to decentralization and network partitions, posts in a timeline form a
causal graph with only partial order, which we promote to a total order based
on the reputation of authors.
The consensus order is fundamental to detect conflicting operations, such as
likes with insufficient reputation (akin to Bitcoin's double spending).
%
Our second contribution is to allow users to create diversified forums of
interest (instead of a singleton blockchain), each counting as an independent
timeline with its own subjective consensus etiquette.
%
Our third contribution is to support content removal without compromising the
integrity of the decentralized blockchain.
Users have the power to revoke posts with dislikes, and peers are forced to
remove payloads, only forwarding associated blockchain metadata.
%
We integrated the proposed consensus algorithm into \FC~\cite{fcs.sbseg20}, a
practical peer-to-peer (P2P) content dissemination protocol that provides
strong eventual consistency~\cite{p2p.sec}.
%p2p.crdts,
To show the practicability of the consensus mechanism, we simulated months of
activity of a chat channel and years of a newsgroup forum, both extracted from
publicly available Internet archives.

@@

In our system, users can spend tokens named \reps to post and rate content in
forums:
a \code{post} initially penalizes authors until it consolidates and counts
positively;
a \code{like} is a positive feedback that helps subscribers to distinguish
content amid excess;
a \code{dislike} is a negative feedback that revokes content when crossing a
threshold.
Table~\ref{fig.general} summarizes the general reputation rules and their
goals.
To prevent Sybils, users with no \reps cannot operate under these rules,
requiring a welcoming like from any other user already in the system.
The fact that likes are zero-sum operations, which only transfer reputation,
eliminates the incentives from malicious users to invite Sybils into the
system.
The only way to generate new \reps is to post content that other users
tolerate, which demands non-trivial work resistant to automation.%
\footnote{
    Note that although human and AI content may become indistinguishable, their
    frequency and relevance in forums is still subject to human evaluation in
    our proposal.
}

@@

Bitcoin employs proof-of-work to mitigate Sybil attacks.
However, CPU or other extrinsic resources are not evenly distributed among
humans, specially in communications using battery-powered devices.
%
Instead, considering the context of public forums, we propose to take advantage
of the human authoring ability as an intrinsic resource.
Creating new content is hard and takes time, but is comparatively easy to
verify and rate.
Therefore, in order to impose scarcity, we determine that only content
authoring generates \reps, while likes and dislikes only transfer \reps between
users.
%but limited by periods of time to prevent excess (e.g., once in a day per user).
%Additionally,

@@

Figure~\ref{fig.reps}.A illustrates the evolution of a forum DAG in the
presence of a network partition:
A common prefix has signed posts from users $a$, $b$, and $c$, when they were
still all connected.
%
We assume that within the prefix, users $a$ and $b$ have contributed with
better content and therefore have more reputation combined than $c$ has alone
(i.e., $8+5>3$).
We will use this fact to order branches in the consensus algorithm.
%
Then, user $c$ disconnects from $a$ and $b$, and evolves with new users $x$ and
$y$ in \code{branch-1}.
In the meantime, $a$ and $b$ participate with new posts in \code{branch-2}.
After some time, the partitions reconnect and new posts from $x$, $b$ and $c$
merge the branches now with a common suffix.

@@

Note that even non-malicious network partitions may experience conflicts when
merging branches.
For instance, dislike operations in one branch to past posts in the common
prefix could possibly block posts from users in the other branch, depending on
how they posts are ordered.
Therefore, we need to order branches with a global consensus, such that we can
validate operations consistently across the network.
%
Note also that, to reach consensus, we should preferably rely on the maximum
information that all branches agree on, which is exactly the posts in the
common prefix.

Based on the reputation of users in the common prefix, Figure~\ref{fig.reps}.B
indicates the consensus order between posts in the forum DAG:
    the common prefix comes first,
    then all posts in \code{branch-2},
    then all posts in \code{branch-1},
    then the common suffix.
%
The \code{branch-2} with users $a$ and $b$ takes priority over \code{branch-1}
with user $c$ because, before the forking point, $a$ and $b$ have more
reputation than $c$, $x$, and $y$ combined.

@@

There are some other relevant considerations about forks and merges:
%
Peers that first receive branches with less reputation need to reorder all
posts after the forking point.
This might involve removing content in the end-user software.
This behavior is similar to Bitcoin's blockchain reorganization, when peers
detect new longest chains. % and disconsiders old blocks.
%
Likewise, peers that first saw branches with more reputation just need to
append the other branch, with no reordering at all.
%and do not need to recompute anything.
This behavior is expected to happen in the majority of the connected network.
%
Note that unlike Bitcoin, forks are not only permitted but encouraged due to
the local-first software principle~\cite{p2p.local}, in which networked
applications can work locally while offline.
However, the longer a peer remains disconnected, the more conflicting
operations it may see, and the higher are the chances of posts reordering when
rejoining.

In summary, it is important to remark that
    (i) branches always introduce conflicts, which
    (ii) require consensus to order operations, which
    (iii) is based on a common prefix reputation,
    (iv) leading to a global and deterministic result.
%
Note that the consensus order exists only to account reputation and verify
operations, and is only a view of the primary DAG structure of public forums.

@@

As a counterpoint, maybe users $a$ and $b$ have abandoned the forum for months,
and thus \code{branch-1} is actually legitimate.
In this case, users $a$ and $b$ might be the ones trying to take over the
forum and would succeed, since they control the majority of \reps in the common
prefix.
Yet another possibility is that both branches are legitimate but became
disconnected for a long period of time.
In any case, it is unacceptable that a very old remote branch affects a long
active forum.

@@

In summary, the rules to merge a branch \code{j} from a remote machine into a
branch \code{i} from a local machine are as follows:
\begin{itemize}
    \item \code{i} is first if it crosses the activity threshold of
          \emph{7 days} or \emph{100 posts}, regardless of \code{j}.
    \item \code{i} or \code{j} is first, whichever has more reputation in the
          common prefix, as discussed in Section~\ref{sec.design.basic}.
    \item otherwise, branches use the arbitrary criteria of lexicographical
          order of the post hashes immediately after the common prefix.
\end{itemize}

@@

Because the protocol relies on tamper-proof DAGs, and because each peer self
verifies the posts as they synchronize, the protocol becomes tolerant to
Byzantine faults~\cite{lamport.byz}.
%
Note that no information coming from any peer is trusted a priori.
%
For instance, a malicious peer that tries to synchronize branches with erratic
data can be detected on the first inconsistent post.
This allows the correct peer to close the connection immediately and blacklist
the malicious peer from future connections.

Therefore, although Byzantine peers may still perform some sort of denial of
service attacks in the network, they cannot modify the DAG structure, nor
create arbitrary content, nor impede that two correct nodes synchronize
directly.

@@

\FC~\cite{fcs.sbseg20} is an unstructured P2P topic-based publish-subscribe
protocol, in which each \emph{chain} is a replicated Merkle-DAG representing
the causal relationships between posts.
%
The protocol operation is typical of publish-subscribe systems: an author
publishes a post to a chain, which subscribed users eventually receive.
% by synchronizing DAGs.

@@

In order to support content moderation and mitigate abuse, we integrated the
proposed reputation and consensus mechanism of Section~\ref{sec.design} with
\FC.
Table~\ref{fig.rules} details the concrete reputation rules we conceived for
\FC, which is compatible with the general rules of Table~\ref{fig.general}, and
which are discussed as follows through an example.
Authors have to sign posts in order to be accounted by the reputation system
and operate in the chains.
We start by creating an identity whose public key is assigned as the pioneer in
a chain named \emph{forum}:
